Visualizing Conditional Dependencies

### 2018-01-31 paper/sections/catKcatV.tex
full draft of intro

\section{Qualitative Variables Conditioned on Qualitative Distributions}

@@

The UpSet tool is designed to display a much larger number of set intersections and the aggregate information of quantitative attributes and set membership. \cite{lex_upset:_2014} As shown in figure~\ref{fig:upsetfig}, the Upset tool provides linked set and element views. The set view provides the intersections and their
aggregates, the frequencies of each set and intersection, and the aggregate
statistics of the associated quantitative variables, while the element view shows
selected elements and information about their set membership. The Upset tool
also facilitates comparison of filtered sets via scatterplot, which acts as a
visualization of the attributes of the data conditioned on membership in the
filtered set.

>>>>

\section{Categorical Variables Conditioned on Categorical Distributions}

@@

The UpSet tool is designed to display a much larger number of set intersections and the aggregate information of quantitative attributes and set membership. \cite{lex_upset:_2014} As shown in figure~\ref{fig:upsetfig}, the Upset tool provides linked set and element views. The set view provides the intersections and their
aggregates, the frequencies of each set and intersection, and the aggregate
statistics of the associated quantitative variables, while the element view shows
selected elements and information about their set membership. The set view yields the distributions of each of the attributes. Far example the selected row (in the gray box) are the distributions of each quantative variable in the comedy set. Every row underneath is the distribution of that quantative variable restricted to movies in the intersection of the selected sets. For example, the distribution of ratings of movies that are both comedies and dramas. The Upset tool
also facilitates comparison of filtered sets via scatterplot, which acts as a
visualization of the attributes of the data conditioned on membership in the
filtered set.

### 2018-01-31 paper/sections/quantKquantV.tex
full draft of intro

While box plots were designed to visualize discrete observations, often there
are no discrete observations and instead the `observation` is a timeseries
curve (where the connections are as important as the individual points) or
spatial region. Understanding the distributions of functional observations
often relies first ordering the functions such that there's some measure of how close these functions are to each other--functional depth\cite{febrero_functional_2007},
bivariate score depth\cite{rob_j._hyndman_rainbow_2010}, and
bivariate kernel density estimation\cite{scott_multivariate_1992} are some examples. This
ordering then yields functional analogs to median and other percentile
metrics. For small values, the ordered curves can then be colored based on the
ordering metric \cite{rob_j._hyndman_rainbow_2010}, but this method is unwieldy for larger
ensembles.

\begin{figure}
  \includegraphics[width=.75\linewidth]{contour_weather}
  \caption{From \cite{whitaker_contour_2013}, the contour box plot method is used to show uncertainty in temperature ensembles while maintaining the temporal dimension of the data.}
  \label{fig:countour}
\end{figure}
Functional bag, box, and HDR plots \cite{rob_j._hyndman_rainbow_2010, sun_functional_2011} tend to be good at visualizing specific types of outliers, but are somewhat limited to only one functional distribution. A very common task is to compare ensembles, each a function of the same timeseries or geographic region. Contour box plots\cite{whitaker_contour_2013} aim to capture the variability of the
uncertainty via a band depth ordering metric. Each ensemble members band depth
is computed as sum of the probabilities that the observations in any given
ensemble fall within the max-min envelope defined by any two other
ensembles. The bands are then sorted by band depth such that the median is the
ensemble with about 50\% of its members whiten all envelopes formed by other
bands (so most centered). Outer bands are chosen according to the task at
hand, which in this paper is temperature visualizations. The authors argue that contour box plots are an improvement over spaghetti plots \cite{luo_visualizing_2003} because the spaghetti plots get nosy as the number of plots increase and so it's hard to tease out specific patterns. Instead in
Figure~\ref{fig:countour}, the authors remove most of the bands and instead visualize an outlying envelope(light gray) and a more central envelope(dark gray). It retains much of the information of
the spaghetti, but removes the visual noise of the lines.

>>>>

While box plots were designed to visualize discrete observations, often there
are no discrete observations and instead the `observation` is a functional densities, drawn from a continuous sample of observations \cite{ramsay_functional_2006, muller_functional_2006}.
One such example is timeseries curve (where the connections are as important as the individual points) or
spatial region. Understanding the distributions of functional observations
often relies first ordering the functions such that there's some measure of how close these functions are to each other--depth\cite{febrero_functional_2007},
bivariate score depth\cite{rob_j._hyndman_rainbow_2010}, and
bivariate kernel density estimation\cite{scott_multivariate_1992} are some examples. This
ordering then yields functional analogs to median and other percentile
metrics. For small values, the ordered curves can then be colored based on the
ordering metric \cite{rob_j._hyndman_rainbow_2010}, but this method is unwieldy for larger
ensembles.

\begin{figure}
  \includegraphics[width=.75\linewidth]{contour_weather}
  \caption{From \cite{whitaker_contour_2013}, the contour box plot method is used to show uncertainty in temperature ensembles while maintaining the trajectory of of the data.}
  \label{fig:countour}
\end{figure}
Functional bag, box, and HDR plots \cite{rob_j._hyndman_rainbow_2010, sun_functional_2011} tend to be good at visualizing specific types of outliers, but are somewhat limited to only one functional distribution. A very common task is to compare ensembles, each a function of the same timeseries or geographic region. Contour box plots\cite{whitaker_contour_2013} aim to capture the variability of the
uncertainty via a band depth ordering metric. Each ensemble members band depth
is computed as sum of the probabilities that the observations in any given
ensemble fall within the max-min envelope defined by any two other
ensembles. The bands are then sorted by band depth such that the median is the
ensemble with about 50\% of its members whiten all envelopes formed by other
bands (so most centered). Outer bands are chosen according to the task at
hand, which in this paper is temperature visualizations. The authors argue that contour box plots are an improvement over spaghetti plots \cite{luo_visualizing_2003} because the spaghetti plots get nosy as the number of plots increase and so it's hard to tease out specific patterns. Instead in
Figure~\ref{fig:countour}, the authors remove most of the bands and instead visualize an outlying envelope(light gray) and a more central envelope(dark gray). It retains much of the information of
the spaghetti, but removes the visual noise of the lines.
